---
title: "map_wa"
author: "Phil Neff, UWCHR"
date: "8/06/2021"
output: html_document
---

# WA state I-213 apprehensions

This notebook performs preliminary descriptive analysis and mapping of ICE/CBP apprehensions in Washington state, based on a collection of I-213 forms obtained by the University of Washington Center for Human Rights (UWCHR) via a Freedom of Information Act (FOIA) lawsuit against the US Department of Homeland Security.

Data used in this analysis were scraped from PDF forms and cleaned in a separate private repository. I-213 narratives and other potentially sensitive fields have been excluded from analysis. Data scraping and cleaning code available for review upon request.

```{r setup, echo=TRUE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, lubridate, here, skimr, yaml, ggplot2, ggalt, ggsflabel,
               sf, rnaturalearth, rnaturalearthdata, maps, tools, RColorBrewer
	)

inputfile <- here::here('map', 'input', 'uw-chr-i213-public.csv.gz')

i213 <- read_delim(inputfile, delim = "|", 
    col_types = cols(
        .default                     = col_character(),
        source                       = col_factor(),
        sex                          = col_factor(),
        year                         = col_double(),
        month                        = col_double(),
        day                          = col_double(),
        hour                         = col_double(),
        minute                       = col_double(),
        age                          = col_double(),
        accompanied_juvenile_flag    = col_double(),
        unaccompanied_juvenile_flag  = col_double(),
        custody_redetermination_flag = col_double())
)

i213 <- i213 %>% 
  filter(source == 'ice',
         state == 'WA',
         year >= '2019')

priority_counties <- read_yaml('../../shared/hand/priority_counties.yaml')

priority_counties <- gsub(" County", "", priority_counties)
priority_counties <- tolower(priority_counties)

priority_plus <- append(priority_counties, 'whatcom')

i213$date <- as.Date(i213$date)

min_date <- min(i213$date, na.rm = TRUE)
max_date <- max(i213$date, na.rm = TRUE)

```

## Mapping I-213 apprehensions by `at_near` location

Most (but not all) I-213 forms include a field labeled "At/Near" that describes the location of the apprehension, usually as a "city, state" pair, but occasionally using other values such as the name of a jail or prison, or agency code for points of entry, such as "SEA" (SeaTac Airport) or "PHY" (Pacific Highway in Blaine, WA). Where "At/Near" values are ambiguous or missing, UWCHR has confirmed apprehension location based on I-213 narratives. After cleaning, locations are assigned latitude/longitude values using Google Maps API. 

```{r map_points, echo=TRUE, warning=FALSE}

world <- ne_countries(scale = "medium", returnclass = "sf")

states <- st_as_sf(map("state", crs = 4326, plot = FALSE, fill = TRUE))

states <- cbind(states, st_coordinates(st_point_on_surface(states)))
states$ID <- toTitleCase(as.character(states$ID))

wa <- states %>% 
  filter(ID == 'Washington')

counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))

counties <- counties %>% 
  separate(col = 'ID', into = c("state", "county"), sep = ",", remove = FALSE) %>% 
  filter(state == 'washington')

counties$area <- as.numeric(st_area(counties))

sites <- i213 %>% select(lon, lat) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("lon", "lat"),
                   crs = 4326, agr = "constant")

wa_coords <- coord_sf(xlim = c(-125, -116.5), ylim = c(45, 49.5), expand = FALSE, crs = 4326)

st_crs(wa) <- 4326
st_crs(wa$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

ggplot(data = wa) +
    geom_sf() +
    geom_sf(data = wa, fill = NA) + 
    geom_sf(data = counties, fill = NA) + 
    geom_sf(data = sites, size = 2, shape = 23, fill = "red") +
    wa_coords +
    ggtitle('I-213 forms by "at_near" latitude/longitude')

```

Note distribution of arrests throughout the state, with larger clusters along the northern border in Whatcom County and in SeaTac, locaiton of both the Seattle-Tacoma International Airport and SeaTac Federal Detention Center. 

```{r map_sites_magnitude, echo=TRUE, message=FALSE, warning=FALSE}

data_at_near <- i213 %>% 
  filter(year >= '2019') %>% 
  group_by(at_near, lon, lat) %>% 
  summarise(n = n()) %>% 
  na.omit()

data_at_near <- st_as_sf(data_at_near, coords = c("lon", "lat"),
                  crs = 4326, agr = "constant")

data_at_near$rank <- rank(data_at_near$n)

p1 <- ggplot(data = wa) +
    geom_sf() +
    geom_sf(data = wa, fill = NA) +
    geom_sf(data = counties, fill = NA) + 
    geom_sf(data = data_at_near, aes(size=n), shape = 21, fill = "red", alpha=.5) +
    wa_coords +
    labs(x = NULL, y = NULL, 
    title = "Count of I-213 apprehensions by `at_near` location",
    subtitle = paste(min_date, max_date, sep = ' to '),
    caption = "UW Center for Human Rights")

p1

ggsave('wa-i213-count.png',
       plot = p1,
       device = 'png',
       path = '../output',
       scale = 1,
       dpi = 300)

p2 <- ggplot(data = wa) +
    geom_sf() +
    geom_sf(data = wa, fill = NA) +
    geom_sf(data = counties, fill = NA) + 
    geom_sf(data = data_at_near, aes(size=n), shape = 21, fill = "red", alpha=.5) +
    ggsflabel::geom_sf_label_repel(data = subset(data_at_near, rank > (nrow(data_at_near) - 5)),
                        aes(label = paste(at_near, n, sep=': '),
                            background = NA),
                        force = 50,
                        nudge_x = -.33,
                        nudge_y = .33) +
    wa_coords +
    labs(x = NULL, y = NULL, 
    title = "Count of I-213 apprehensions by `at_near` location",
    subtitle = paste(min_date, max_date, sep = ' to '),
    caption = "UW Center for Human Rights")

p2

ggsave('wa-i213-count-labeled.png',
       plot = p2,
       device = 'png',
       path = '../output',
       scale = 1,
       dpi = 300)

```

Here we sum total arrests by county, and also generate a simple diagram of priority counties selected for UWCHR's "Immigrant Rights Observatory" project.

```{r map_counties_total, echo=TRUE, warning=FALSE, message=FALSE}

data <- i213 %>% 
  group_by(county, state) %>% 
  summarise(n = n()) %>% 
  na.omit()

data$state <- data$state %>% 
  str_replace_all(c("WA" = "washington", "OR" = "oregon", 'ID' = 'idaho', 'AK' = 'alaska'))

data$county <- data$county %>% 
  str_replace_all(c(" County" = "")) %>% 
  tolower()

data$ID <- paste(data$state, data$county, sep=',')

counties <- left_join(counties, data, by=c("ID", "state", 'county'))

study_counties <- counties %>% 
  filter(county %in% priority_counties)

m1 <- ggplot(data = wa) +
    geom_sf() +
    geom_sf(data = wa, fill = NA) + 
    geom_sf(data = counties, aes(fill = n)) +
    ggsflabel::geom_sf_label_repel(data = study_counties, aes(label = str_to_title(county)), 
                        nudge_x = -.33,
                        nudge_y = .33,
                        force = 75) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4, na.value = gray(.9)) +
    wa_coords +
    labs(x = NULL, y = NULL, 
       title = "Total I-213 apprehensions by county",
       subtitle = paste(min_date, max_date, sep = ' to '),
       caption = "UW Center for Human Rights")

m1

ggsave('wa-i213-county-total.png',
       plot = m1,
       device = 'png',
       path = '../output',
       scale = 1,
       dpi = 300)

m2 <- ggplot(data = wa) +
    geom_sf() +
    geom_sf(data = wa, fill = NA) + 
    geom_sf(data = counties, fill = NA) +
    geom_sf(data = study_counties, aes(fill = 'priority')) +
    geom_sf_label_repel(data = study_counties, aes(label = str_to_title(county), background = NA), force = 0) +
    wa_coords +
    labs(x = NULL, y = NULL, fill = NULL,
       title = "Immigrant Rights Observatory priority counties",
       caption = "UW Center for Human Rights") +
    annotate("text", x = 4, y = 25, label = "Some text") +
    theme(legend.position = "none") + scale_fill_manual(values=c("#b7a57a"))

m2 

ggsave('wa-priority-counties.png',
       plot = m2,
       device = 'png',
       path = '../output',
       scale = 1,
       dpi = 300)

```

Here we calculate and map rate of I-213 apprehensions per capita, using ACS 5 year 2014-2018 population estimates.

```{r map_counties_per_capita, echo=TRUE, message=FALSE, warning=FALSE}

wa_pop <- read.delim(here::here('map', 'input', 'wa-pop.csv'), sep=',')

wa_total_pop <- wa_pop %>%  
  filter(name == "washington")

wa_total_pop <- wa_total_pop$estimate

wa_pop <- wa_pop %>% 
  filter(name != "washington")

wa_pop$name <- gsub(" county", "", wa_pop$name)
wa_pop$name <- str_trim(wa_pop$name)

wa_pop <- wa_pop %>% 
  separate(., col=name, sep=',',
           remove = FALSE,
           into=c('county', 'state'))

wa_pop$state <- str_trim(wa_pop$state)
wa_pop$county <- str_trim(wa_pop$county)

wa_pop <- wa_pop %>%
  mutate(priority = county %in% priority_counties)

wa_pop <- wa_pop %>%
  mutate(priority_plus = county %in% priority_plus)

wa_pop_prior <- wa_pop %>% 
  filter(priority == TRUE)

wa_pop_prior_plus <- wa_pop %>% 
  filter(priority_plus == TRUE)

wa_pop_prior <- sum(wa_pop_prior$estimate)
wa_pop_prior_plus <- sum(wa_pop_prior_plus$estimate)

prop_pop_prior <- wa_pop_prior / wa_total_pop
wa_pop_prior_plus <- wa_pop_prior_plus / wa_total_pop

# merge and calcuiate n forms per capita

data <- merge(counties, wa_pop, by=c('county', 'state'))

data$i_213_per_capita <- data$n / data$estimate * 100000

# Checking # of forms in priority counties
forms_wa <- data %>%
  filter(state == 'washington')

forms_prior <- data %>% 
  filter(priority == TRUE)

forms_prior_plus <- data %>% 
  filter(priority_plus == TRUE)

wa_forms_prior <- sum(forms_prior$n) / sum(forms_wa$n)

wa_forms_prior_plus <- sum(forms_prior_plus$n) / sum(forms_wa$n)

total_forms_prior <- sum(forms_prior$n) / sum(data$n)

m1 <- ggplot(data = wa) +
    geom_sf() +
    geom_sf(data = wa, fill = NA) + 
    geom_sf(data = data, aes(fill = i_213_per_capita)) +
    wa_coords +
    ggsflabel::geom_sf_label_repel(data = study_counties, aes(label = str_to_title(county), background = NA), force = 100) +
    labs(x = NULL, y = NULL, 
       title = "Per capita I-213 apprehensions by county",
       subtitle = paste(min_date, max_date, sep = ' to '),
       caption = "UW Center for Human Rights\n2014-2018 population estimate via ACS") +
  guides(fill=guide_colorbar(title="Arrests per\n100,000")) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4, na.value = gray(.9))

m1

ggsave('wa-i213-county-per-capita.png',
       plot = m1,
       device = 'png',
       path = '../output',
       scale = 1,
       dpi = 300)

```

Table of apprehensions per capita:

```{r table_counties_per_capita, echo=TRUE, message=FALSE, warning=FALSE}

to_print <- data %>%
  select(c('county', 'n', 'estimate', 'i_213_per_capita')) %>% 
  mutate(county = str_to_title(county))
  
st_geometry(to_print) <- NULL

total_arrests <- sum(to_print$n, na.rm=TRUE)
wa_per_capita <- total_arrests / wa_total_pop * 100000

to_print <- to_print %>% 
  add_row(county = 'Washington state total',
          estimate = wa_total_pop,
          n = total_arrests,
          i_213_per_capita = wa_per_capita)

to_print <- to_print %>% 
  arrange(desc(i_213_per_capita))

col_names <- c('County', 'Total apprehensions', 'Population','Apprehensions per capita')

caption <- paste("WA State I-213 apprehensions per capita\n", min_date, " to ", max_date, sep='')
caption_html <- paste("<b>WA State I-213 apprehensions per capita</b><br><i>", min_date, " to ", max_date, sep='')

write(knitr::kable(to_print, caption = caption_html, col.names = col_names, digits=4, 'html'), '../output/i213-per-capita.html')
print(knitr::kable(to_print, caption = caption, col.names = col_names, digits=4))
```
