---
title: JSIS Task Force analysis
author:
- "[Phil Neff](https://github.com/philneff)"
date: 10 January, 2021
output:
    html_document:
        html_preview: true
        toc: true
        toc_depth: 3
        toc_float: true
        code_folding: hide
---

```{r import, message = FALSE, warning = FALSE}

pacman::p_load(tidyverse, lubridate, here, skimr, yaml, ggplot2, ggalt, gplots,
               sf, rnaturalearth, rnaturalearthdata, maps, tools, RColorBrewer
               )

inputfile <- here::here("write/input/uw-chr-i213-public.csv.gz")

i213 <- read_delim(
    inputfile, delim = "|", 
    col_types = cols(
        # .default                     = col_character(),
        source                       = col_character(),
        sex                          = col_character(),
        cmplxn                       = col_character(),
        country_of_citizenship       = col_factor(),
        year                         = col_double(),
        month                        = col_double(),
        day                          = col_double(),
        hour                         = col_double(),
        minute                       = col_double(),
        fy                           = col_double(),
        age                          = col_double(),
        accompanied_juvenile_flag    = col_double(),
        unaccompanied_juvenile_flag  = col_double(),
        custody_redetermination_flag = col_double()
))

i213 <- i213 %>%
  mutate(across(where(is.character), ~ na_if(.,"")))

# Only looking at records obtained via ICE, which include Border Patrol/CBP apprehensions
i213 <- i213 %>%
    filter(source != "cbp")

```


## Data overview

```{r skim, message = FALSE, warning = FALSE}

i213 %>% skimr::skim()

```

```{r state}

state <- i213 %>%
   count(state, sort = TRUE)

state

i213 <- i213 %>%
    filter(state %in% c("OR", "WA"))

```

```{r year, message = FALSE, warning = FALSE}


ann <- i213 %>%
    filter(state %in% c("OR", "WA")) %>% 
    count(state, year) %>%
    pivot_wider(names_from = state, values_from = n,
                values_fill = list(n = 0)) %>%
    arrange(year)

ann

```

```{r weekday, message = FALSE, warning = FALSE}

day_of_week <- function(year, month, day) {
    dt <- as.Date(paste(year, month, day, sep = "-"))
    wd <- weekdays(dt, abbreviate = TRUE)
    wd <- as.factor(wd)
    fct_relevel(wd, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
}

wkdy <- i213 %>%
    filter(state %in% c("OR", "WA")) %>% 
    filter(!is.na(day)) %>%
    mutate(weekday = day_of_week(year, month, day)) %>%
    count(state, weekday) %>%
    pivot_wider(names_from = state, values_from = n)

wkdy

```

```{r year_month}

# Change to include year?

mth <- i213 %>%
    filter(state %in% c("OR", "WA")) %>% 
    filter(!is.na(month)) %>%
    count(state, month) %>%
    pivot_wider(names_from = month, values_from = n)

mth

```


```{r time_of_day}

time <- i213 %>%
    filter(state %in% c("OR", "WA")) %>% 
    filter(!is.na(hour)) %>%
    count(state, hour) %>%
    pivot_wider(names_from = hour, values_from = n)

time

```

## Demographics: Overview

"Complexion" values are decoded following ["The Border's Long Shadow"](https://www.aclumich.org/en/publications/borders-long-shadow) (ACLU of Michigan, 2021).

```{r demographics, message = FALSE, warning = FALSE}

i213 %>% 
    filter(state %in% c("OR", "WA")) %>% 
  group_by(state, country_of_citizenship) %>% 
  count(country_of_citizenship, sort=TRUE) %>% 
    pivot_wider(names_from = state, values_from = n)
  
i213 %>%
    filter(state %in% c("OR", "WA")) %>% 
    filter(!is.na(age), !is.na(sex)) %>%
    mutate(age_group = cut(age,
                           breaks = c(0, 18, 25, 40, 60, Inf),
                           right = FALSE)) %>%
    group_by(state, sex, age_group) %>% 
    count(sex, age_group) %>%
    pivot_wider(names_from = sex, values_from = n)

good_values <- c("ALBINO",
                 "BLACK",
                 "DARK BROWN",
                 "DARK",
                 "FAIR",
                 "LIGHT BROWN",
                 "LIGHT",
                 "MEDIUM BROWN",
                 "MEDIUM",
                 "OLIVE",
                 "RUDDY",
                 "SALLOW",
                 "YELLOW")

i213 <- i213 %>%
          mutate(cmplxn = case_when(
            cmplxn %in% good_values ~ cmplxn,
            is.na(cmplxn) ~ NA_character_,
            TRUE ~ "OTHER"))

i213 %>% 
  filter(state %in% c("OR", "WA")) %>% 
  group_by(state, cmplxn) %>% 
  count(cmplxn, sort = TRUE)

```

```{r at_near, message = FALSE, warning = FALSE}

i213 %>% count(at_near, sort = TRUE) %>% 
   filter(n >= 20)

```


```{r loc_code, message = FALSE, warning = FALSE}

i213 %>% count(location_code, sort = TRUE) %>% 
   filter(n >= 20)

i213$office <- str_match(i213$location_code, "SEA|POO|BLW|SPW")

loc_codes <- i213 %>% 
  group_by(location_code) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  na.omit()

loc_codes_county <- i213 %>% 
  group_by(location_code) %>% 
  summarize(n_counties = n_distinct(county),
            n_state = n_distinct(state)) %>% 
  arrange(desc(n_counties)) %>% 
  na.omit()

top <- unlist(loc_codes[1:20, 'location_code'])

```

## "Method of Location/Apprehension"

```{r method_location_apprehension_all, message = FALSE, warning = FALSE}

i213 %>%
  filter(!is.na(method_location_apprehension)) %>% 
  group_by(state, method_location_apprehension) %>% 
  count(method_location_apprehension, sort = TRUE)

```

```{r method_location_apprehension, message=FALSE, warning=FALSE}

i213 <- i213 %>% 
  mutate(method_short = case_when(
           str_detect(method_location_apprehension, "PB") ~ "PB",
           str_detect(method_location_apprehension, "CFD") ~ "CFD",
           str_detect(method_location_apprehension, "CST") ~ "CST",
           str_detect(method_location_apprehension, "CLC") ~ "CLC",
           str_detect(method_location_apprehension, "NCA") ~ "NCA",
           str_detect(method_location_apprehension, "LEA") ~ "LEA",
           str_detect(method_location_apprehension, "OA") ~ "OA",
           str_detect(method_location_apprehension, "OTF") ~ "OTF",
           str_detect(method_location_apprehension, "TCB") ~ "TCB",
           str_detect(method_location_apprehension, "ISP") ~ "ISP",
           str_detect(method_location_apprehension, "L") ~ "L",
           str_detect(method_location_apprehension, "O|0") ~ "O",
           is.na(method_location_apprehension) ~ NA_character_,
            TRUE ~ NA_character_),
         method_verbose = case_when(
           str_detect(method_location_apprehension, "PB") ~ "Patrol Border",
           str_detect(method_location_apprehension, "CFD") ~ "CAP Federal",
           str_detect(method_location_apprehension, "CST") ~ "CAP State",
           str_detect(method_location_apprehension, "CLC") ~ "CAP Local",
           str_detect(method_location_apprehension, "NCA") ~ "Non-Custodial Arrest",
           str_detect(method_location_apprehension, "LEA") ~ "Law Enforcement Assist",
           str_detect(method_location_apprehension, "OA") ~ "Other Agency",
           str_detect(method_location_apprehension, "OTF") ~ "Task Force",
           str_detect(method_location_apprehension, "TCB") ~ "Transportation Check",
           str_detect(method_location_apprehension, "ISP") ~ "Inspection",
           str_detect(method_location_apprehension, "L") ~ "Located",
           str_detect(method_location_apprehension, "O|0") ~ "Other",
           is.na(method_location_apprehension) ~ NA_character_,
            TRUE ~ NA_character_),
         method_int_border = case_when(
           str_detect(method_location_apprehension, "PB") ~ "Border",
           str_detect(method_location_apprehension, "CFD") ~ "Interior",
           str_detect(method_location_apprehension, "CST") ~ "Interior",
           str_detect(method_location_apprehension, "CLC") ~ "Interior",
           str_detect(method_location_apprehension, "NCA") ~ "Interior",
           str_detect(method_location_apprehension, "LEA") ~ "Interior",
           str_detect(method_location_apprehension, "OA") ~ "Interior",
           str_detect(method_location_apprehension, "OTF") ~ "Interior",
           str_detect(method_location_apprehension, "TCB") ~ "Interior",
           str_detect(method_location_apprehension, "ISP") ~ "Border",
           str_detect(method_location_apprehension, "L") ~ "Interior",
           str_detect(method_location_apprehension, "O|0") ~ "Other",
           is.na(method_location_apprehension) ~ NA_character_,
            TRUE ~ NA_character_))

i213$date <- as.Date(i213$date)

data <- i213 %>%
  select(date, state, source, method_location_apprehension, method_short, method_verbose, method_int_border, office, county) %>%
  filter(!is.na(state),
         date >= '2019-01-01',
         source == 'ice')

data$month <- as.Date(cut(data$date, breaks='month'))

p1 <- data %>% 
  ggplot(aes(x = month)) +
  geom_bar() +
  scale_x_date(labels = scales::date_format("%b"),
               breaks = "1 month") +
  facet_wrap(~method_short) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p1

p2 <- data %>%
  ggplot(aes(x = method_short)) +
  geom_bar(aes(fill= method_short)) +
  facet_wrap(~office) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p2

priority_counties <- read_yaml('../../shared/hand/priority_counties.yaml')
priority_counties <- append(priority_counties, 'Whatcom County')

p3 <- data %>% 
  filter(county %in% priority_counties) %>% 
  ggplot(aes(x = county, fill= method_verbose)) +
  geom_bar(position = "fill") +
  scale_y_continuous() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p3

p4 <- data %>% 
  filter(!county %in% priority_counties) %>%
  ggplot(aes(x = county, fill= method_verbose)) +
  geom_bar(position = "fill") +
  scale_y_continuous() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p4

p5 <- data %>% 
  filter(county %in% priority_counties) %>%
  ggplot(aes(x = county, fill= method_int_border)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p5

```

```{r method_location_apprehension_city_wa, message = FALSE, warning = FALSE}

data <- i213 %>%
  select(date, state, source, method_location_apprehension, method_short, method_verbose, kww_in_force, office, county, at_near) %>%
  filter(!is.na(state),
         state == 'WA',
         date >= '2019-01-01',
         source == 'ice')

at_near_clc <- data %>% group_by(at_near, method_short) %>% 
  summarise(n = n()) %>% 
  mutate(percent = n/sum(n)) %>% 
  filter(method_short == 'CLC')

at_near_clc <- at_near_clc[order(-at_near_clc$n),]

head(at_near_clc, 10)

at_near_assist <- data %>% group_by(at_near, method_short) %>% 
  summarise(n = n()) %>% 
  mutate(percent = n/sum(n)) %>% 
  filter(method_short %in% c('LEA', 'OA', 'OTF'))

at_near_assist <- at_near_assist[order(-at_near_assist$n),]

```

```{r method_location_apprehension_city_or, message = FALSE, warning = FALSE}

data <- i213 %>%
  select(date, state, source, method_location_apprehension, method_short, method_verbose, kww_in_force, office, county, at_near) %>%
  filter(!is.na(state),
         state == 'OR',
         date >= '2019-01-01',
         source == 'ice')

at_near_clc <- data %>% group_by(at_near, method_short) %>% 
  summarise(n = n()) %>% 
  mutate(percent = n/sum(n)) %>% 
  filter(method_short == 'CLC')

at_near_clc <- at_near_clc[order(-at_near_clc$n),]

head(at_near_clc, 10)

at_near_assist <- data %>% group_by(at_near, method_short) %>% 
  summarise(n = n()) %>% 
  mutate(percent = n/sum(n)) %>% 
  filter(method_short %in% c('LEA', 'OA', 'OTF'))

at_near_assist <- at_near_assist[order(-at_near_assist$n),]

```

```{r method_location_apprehension_or, message = FALSE, warning = FALSE}

data <- i213 %>%
  select(date, state, source, method_location_apprehension, method_short, method_verbose, kww_in_force, office, county) %>%
  filter(!is.na(state),
         date >= '2019-01-01',
         source == 'ice')

data$month <- as.Date(cut(data$date, breaks='month'))

p1 <- data %>% 
  filter(!is.na(state),
         state == 'OR') %>% 
  ggplot(aes(x = month)) +
  geom_bar() +
  scale_x_date(labels = scales::date_format("%b"),
               breaks = "1 month") +
  facet_wrap(~method_short) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p1

p2 <- data %>%
    filter(!is.na(state),
         state == 'OR') %>% 
  ggplot(aes(x = method_short)) +
  geom_bar(aes(fill= method_short)) +
  facet_wrap(~office) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p2

p3 <- data %>%
    filter(!is.na(state),
         state == 'OR') %>% 
  ggplot(aes(x = county, fill= method_verbose)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p3

p4 <- data %>%
    filter(!is.na(state),
         state == 'OR') %>% 
  ggplot(aes(x = county, fill= method_verbose)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p4

p5 <- data %>%
  filter(!is.na(state),
           state %in% c("OR", "WA")) %>% 
  ggplot(aes(x = state, fill= method_verbose)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p5

p6 <- data %>%
    filter(!is.na(state),
           state %in% c("OR", "WA")) %>% 
  ggplot(aes(x = state, fill= method_verbose)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p6


```

### Mapping "Method of Location/Apprehension"

```{r method_location_apprehension_map, message = FALSE, warning = FALSE}

sites <- i213 %>% select(lon, lat, office, method_short) %>% 
  na.omit()

sites <- st_as_sf(sites, coords = c("lon", "lat"),
                   crs = 4326, agr = "constant")

sites <- cbind(sites, st_coordinates(sites))

st_crs(states) <- 4326
st_crs(states$geom) <- 4326
st_crs(sites) <- 4326
st_crs(sites$geometry) <- 4326

m1 <- ggplot() +
    geom_sf(data = world) +
    geom_sf(data = states, fill = NA) + 
    geom_jitter(data = sites, size = 2, shape = 23, width = .1, height = .1, aes(x = X, y = Y, fill = method_short)) +
    pnw_coords +
    facet_wrap(~method_short)
    ggtitle('I-213 forms by approx "at_near" latitude/longitude')

m1

```

## Demographics: Zooming in

### Complexion by location and enforcement type

We can also examine various demographic characteristics in comparison with the above categories; note below the only "Method of Location/Apprehension" category with a majority of apprehensions reported as being of people with "Fair" or "Light" complexion is "ISP" or "Inspection" at authorized points of entry.

```{r cmplxn_by_loc_meth, message=FALSE, warning=FALSE}

p1 <- i213 %>%
  ggplot(aes(x = cmplxn, fill= cmplxn)) +
  geom_bar() +
  facet_wrap(~office) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p1

p2 <- i213 %>%
  ggplot(aes(x = cmplxn, fill= cmplxn)) +
  geom_bar() +
  facet_wrap(~method_short) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p2

```

### Country of origin

A few views of most common "Country of Origin" categories. All others have <50 records in collection.

```{r country_by_loc_meth, message=FALSE, warning=FALSE}

i213 <- i213 %>% 
  mutate(country_short = case_when(
         str_detect(country_of_citizenship, "MEXICO") ~ "MEXICO",
         str_detect(country_of_citizenship, "GUATEMALA") ~ "GUATEMALA",
         str_detect(country_of_citizenship, "INDIA") ~ "INDIA",
         str_detect(country_of_citizenship, "HONDURAS") ~ "HONDURAS",
         str_detect(country_of_citizenship, "EL SALVADOR") ~ "EL SALVADOR",
         str_detect(country_of_citizenship, "CHINA") ~ "CHINA",
         is.na(country_of_citizenship) ~ NA_character_,
         TRUE ~ "ALL OTHERS"))

p1 <- i213 %>%
  ggplot(aes(x = country_short, fill= country_short)) +
  geom_bar() +
  facet_wrap(~office) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p1

p2 <- i213 %>%
  ggplot(aes(x = country_short, fill= country_short)) +
  geom_bar() +
  facet_wrap(~method_short) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p2

p3 <- i213 %>% 
  filter(county %in% priority_counties) %>% 
  ggplot(aes(x = county, fill= country_short)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p3

i213$month <- as.Date(cut(i213$date, breaks='month'))

p4 <- i213 %>% 
  filter(!is.na(state),
       state == 'WA',
       date >= '2019-01-01',
       source == 'ice') %>% 
  ggplot(aes(x = month)) +
  geom_bar() +
  scale_x_date(labels = scales::date_format("%b"),
               breaks = "1 month") +
  facet_wrap(~country_short) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p4

```

A closer look at Whatcom County data:

```{r whatcom_deep_dive, message=FALSE, warning=FALSE}

p5 <- i213 %>% 
  filter(!is.na(state),
       state == 'WA',
       date >= '2019-01-01',
       source == 'ice',
       county == 'Whatcom County') %>% 
  ggplot(aes(x = country_short, fill= country_short)) +
  geom_bar() +
  facet_wrap(~method_int_border) +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5,
                                   hjust = 1))

p5

p5_data <- i213 %>% 
  filter(!is.na(state),
       state == 'WA',
       date >= '2019-01-01',
       source == 'ice',
       county == 'Whatcom County') %>% 
  group_by(method_verbose, country_of_citizenship) %>%
  summarise(n = n())

# write_csv(p5_data, '../output/whatcom-method-citizenship.csv')

```

## Mapping "mentions_*"

A few passes at mapping records that mention keywords.

Note I-213s with narratives mentioning keyword "courthouse" include high concentration in central WA; see 2019 UWCHR [report on courthouse arrests in WA state](https://jsis.washington.edu/humanrights/2019/10/16/ice-cbp-courthouse-arrests/).

```{r mentions_courthouse_map, message=FALSE, warning=FALSE}

data <- i213 %>% 
  group_by(mentions_courthouse, lon, lat) %>% 
  summarise(n = n()) %>% 
  filter(mentions_courthouse == TRUE) %>% 
  na.omit()

data <- st_as_sf(data, coords = c("lon", "lat"),
                  crs = 4326, agr = "constant")

ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) +
    geom_sf(data = data, aes(size=n), shape = 21, fill = "red", alpha=.5) +
    pnw_coords +
    ggtitle('I-213 forms mentioning "courthouse" by "at_near" latitude/longitude')

```

```{r mentions_juvenile_map, message=FALSE, warning=FALSE}

data <- i213 %>% 
  group_by(mentions_juvenile, lon, lat) %>% 
  summarise(n = n()) %>% 
  filter(mentions_juvenile == TRUE) %>% 
  na.omit()

data <- st_as_sf(data, coords = c("lon", "lat"),
                  crs = 4326, agr = "constant")

ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) +
    geom_sf(data = data, aes(size=n), shape = 21, fill = "red", alpha=.5) +
    pnw_coords +
    ggtitle('I-213 forms mentioning "juvenile" by "at_near" latitude/longitude')

```

```{r mentions_bus_map, message=FALSE, warning=FALSE}

data <- i213 %>% 
  group_by(mentions_bus, lon, lat) %>% 
  summarise(n = n()) %>% 
  filter(mentions_bus == TRUE) %>% 
  na.omit()

data <- st_as_sf(data, coords = c("lon", "lat"),
                  crs = 4326, agr = "constant")

ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) +
    geom_sf(data = data, aes(size=n), shape = 21, fill = "red", alpha=.5) +
    pnw_coords +
    ggtitle('I-213 forms mentioning "bus" by "at_near" latitude/longitude')

```
